### 2022-05-05 talk/talk.qmd
start working on talk

---
title: "Developing Measures of Public Perceptions of Values in Science"
author: |
    | Dan Hicks 
    | Emilio Lobato
    | University of California, Merced
    | *[bitly]*
toc: true
format: 
    revealjs:
        toc-depth: 1
        slide-number: true
        center: false
        <!--margin: 0.0-->
link-citations: true
---

# Controversy and mismatch # {center="true"}

## Mismatch hypothesis

- A mismatch between **public understanding of philosophy of science** and **realities of scientific practice**
- This mismatch contributes to public scientific controversies
- In some cases, this mismatch has been deliberately exploited to manufacture doubt

## John: Climategate vs. transparency

:::: {.columns}
::: {.column}
![\n](img/John.jpg){fig-alt="Photo of Stephen John"}
:::
::: {.column}
*[E]pistemic trust in science may be 'fragile' because based on a false 'folk philosophy of science'.* [@JohnEpistemicTrustEthics2017 6-7]
:::
::::

- Inferring causation from correlation (observational data)
- Refusing (desk rejecting) papers by climate skeptics
- "[I]gnoring recalcitrant or strange data sets" [@JohnEpistemicTrustEthics2017 7]

## Kovaka: Pro-science science denial {.smaller}

:::: {.columns}
::: {.column}
*[P]eople have serious misconceptions about what scientific inquiry ought to look like. Their misconceptions interact with partisan biases to produce denial of human-caused climate change.* [@KovakaClimateChangeDenial2021 1]
:::
::: {.column}
![\n](img/Kovaka.jpg){fig-alt="Photo of Karen Kovaka"}
:::
::::

Misconceptions about

- Objectivity: "scientists as more objective than other people"; value-free ideal
- Scientific method: "a universal, step-by-step scientific method," based on experiments
- Uncertainty and fallibility: Galileo narratives and quest for certainty [@KovakaClimateChangeDenial2021 11-14]

## Nature of science

*[substantial area of research, esp. by education researchers; 
but no scale we found considered role of values in a sophisticated way]

"more than 25 NOS instruments have been developed over the past 50 years" [Neumann citing Lederman 2007]
LiangStudentUnderstandingScience2006 -> Lombrozo et al -> Weisberg et al.
Lombrozo et al version critiqued by Neumann
    - N had science educators map L items to categories based on a lit review of NOS education research (1377, 1384); these don't correspond to L's phil sci-informed categories (1384ff), and don't include issues such as VFI
    - psychometric critique

## *[other empirical research]*

*[Elliott et al; Safford et al]

# Developing a Values in Science Scale (VISS)

## VISS items ##

36 items, 3 each for 12 different topics in the \
science, values, and policy literature

:::: {.columns}
::: {.column}
Aims of science\
Conflicts of interest\
Consensus\
Fact-value distinction\
Fallibilism\
Inductive risk
:::
::: {.column}
Non-subjectivity\
Pluralism\
Scientism\
Standpoint theory\
Technocracy\
Value-free ideal
:::
::::

## Data and methods ##

- Online survey/marketing platform Prolific
- Representative sample of 980 US adults
- Replication of *[elliott et al.]* + VISS + demographics

</br>

- Exploratory factor analysis of VISS items
- Statistical analysis of several findings from *[elliott et al]*

## VISS factors ##

<!--:::: {.columns}
::: {.column width="30%"}
Scientism\
Values in science\
Cynicism\
Power\
Textbook science\
Value-free ideal
:::
::: {.column width="70%"}--->
![\n](../out/02_six_corr_matrix.png){fig-alt="Screenshot of the factor correlation matrix.  The factors are labeled scientism, VIS (values in science), cynicism, power, textbook (science), and VFI (value-free ideal)" height="650"}