### 2022-07-13 readme.Rmd

---
title: "Data and code for \"Values disclosures and trust in science: A replication study\""
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

This repository contains anonymized data and code for the paper "Values disclosures and trust in science: A replication study" by Daniel J. Hicks and Emilio Lobato.

# Data files

Data files are contained in the folder `data`.

The files `data/data.csv` and `data/data.Rds` contain anonymized versions of the original data collected for this replication study.  An overview is included at the bottom of this readme.

The files `data/emad.csv` and `data/emad.Rds` contain versions of the data from Elliot et al. experiment 1, downloaded from <https://figshare.com/articles/dataset/Elliott_McCright_Allen_and_Dietz_Data_SPSS_Stata_/4695793> on 2021-06-09. This dataset was reshaped and variables recoded to facilitate parallel analysis with our original data using the code in script `01_clean.R`. This dataset is made available under a Creative Commons Attribution 4.0 International license.

# Reproducibility

To preserve participant anonymity, raw data (as downloaded from Qualtrics and cleaned by script `01_clean.R`) is not publicly available. Only analysis scripts `02_analysis.R` and `03_dag.R` are reproduced, along with assembling the manuscript PDF.

Either method below will require prior installation of R, the `renv` package, and [Quarto](https://quarto.org/).

## Using make

After cloning and downloading, in Bash or a similar shell run
```
make install
make
```

## Manually

In R, run
```
renv::restore()
```
to create a local (within-project) package library, with the specific versions used in this project.  Then run the scripts `02_analysis.R` and `03_dag.R`, in that order, to run the analyses.  These scripts will write plots and tables to the `out` folder, creating it if necessary.  Optionally, use `rmarkdown::render()` or the "Knit" button in RStudio to generate HTML documents from these scripts, for comparison with the ones included in the repository.  (Be sure to rename the included HTML files first.)

Finally, in the `paper` folder, from Bash or a similar shell run
```
quarto render paper.Qmd --to pdf
```
to generate the PDF version of the manuscript.

# Data overview

```{r}
library(skimr)

dataf = readRDS(file.path('data', 'data.Rds'))
skim(dataf)
```